\documentclass[xcolor=dvipsnames]{beamer}
% Class options include: notes, notesonly, handout, trans,
%                        hidesubsections, shadesubsections,
%                        inrow, blue, red, grey, brown

% Theme for beamer presentation.
\usetheme{Susan}
\usepackage{graphics}
\usepackage{multicol}
\usepackage{url}

% Other themes include: beamerthemebars, beamerthemelined, 
%                       beamerthemetree, beamerthemetreebars  


\section*{Programs with Python)}

\begin{document}

\begin{frame}
\begin{center}{\Huge Building Programs with Python}
\end{center}
\end{frame}

\begin{frame} 
\frametitle{Analyzing Patient Data Part 1}
\begin{enumerate}
\item    Explain what a library is, and what libraries are used for.
\item    Load a Python library and use the things it contains.
\item    Read tabular data from a file into a program.
\item    Assign values to variables.
\item    Select individual values and subsections from data.
\end{enumerate}

\begin{multicols}{2}
\begin{itemize}
\item import numpy
\item numpy.loadtxt(fname=  delimiter=)
\item weight\_kg = 55
\item print
\item weight\_lb = 2.2 * weight\_kg
\item type(data)
\item data.shape
\item data[0,0], data[0:1,0:1]
\item data[0:10:2,1]
\item data[:3,36:]
\end{itemize}
\end{multicols}
\end{frame}

\begin{frame}
\frametitle{Analyzing Patient Data Part 2}
\begin{enumerate}
\item    Perform operations on arrays of data.    
\item    Display simple graphs.  
\end{enumerate}
\begin{multicols}{2}
\begin{itemize}
\item data.mean()
\item data.std()
\item data.mean(axis=0)
\item \%matplotlib inline
\item from matplotlib import pyplot 
\item pyplot.imshow(data)
\item pyplot.show()
\item pyplot.plot(ave\_inflammation)
\item import matplotlib import pyplot as plt
\item plt.subplot(1,3,1)
\item plt.ylabel('average')
\item plt.show()
\end{itemize}
\end{multicols}
\end{frame}


\begin{frame}
\frametitle{Exercise}
Create a single plot showing 1) the mean for each day and 2) the mean + 1 standard deviation for each day and the 3) the mean - 1 standard deviation for each day.
\end{frame}
\end{document}